---
title: "Lecture 3_2 "
author: "Doug nychka"
date: "January 27, 2016"
output: html_document
---


Create a small matrix of numbers to do some examples

```{r}
look<- matrix( 1:12, ncol=4, nrow=3)
look
```
Matrix transpose and dimension
```{r}
dim( look)
t( look)
```
Using cbind to create the matrix column by column
```{r}
look<- cbind( c( 1:3), c(4:6), c(7:9), c(10:12))
```
Convert this matrix to a data frame
```{r}
look2<- data.frame( look)
```
A data frame can have columns of text (aka strings) sometimes very useful. A matrix can not! In this example the whole matrix is converted to text when this is tried.
```{r}
look3<- cbind( animals= c("cat", "dog", "lion"), oldData= look2)
print( look3)
 cbind( c("cat", "dog", "lion"), look)
 ```
The column names have been added to make this print more clearly.
 
Create an NA
```{r}
look[2,3] <- NA
look
look2<- is.na( look)
look2

 look==4 

```

Applying over rows and columns

```{r}
apply( look, 1, "max")

apply( look, 2, "max")

```

Here is an image plot of this matrix
```{r}
image(1:3, 1:4, look, col=rainbow(256))
```

Here is an image plot of this matrix
using the fields function that adds a legend
```{r}
library(fields)
image.plot(1:3, 1:4, look, col=rainbow(256))
```