
More programming tips for data analysis
--

Writing R code often involves adding features that make the results better documented and also making the inputs do the function is easy to use. 

This lecture gives some strategies that are often helpful.

As motivation here is a simple function to fit a quadratic polyomial function to a data set using the **lm** function and return the predicted values. (You need to load the fields package for this to work.)
```{r}
myFit<- function( x,y){
 degree<- 2
 fit<- lm(y~ x + I(x^2) )
 A<- fit$coefficients
 predValues<- A[1] + A[2]*x + A[3]*x^2
return( predValues)
}
```
Note: ```fields.mkpoly``` will return a matrix where each column is a power of x. At this point what lm does and why is still mysterious -- this will be covered when we talk about least squares fitting. 

```{r include=FALSE}
library(dataWorkshop)
```
```{r}
data(AudiA4)
hold<- myFit( AudiA4$mileage, AudiA4$price)
# take a look
plot( AudiA4$mileage, AudiA4$price)
points(AudiA4$mileage,hold, col="seagreen")
```

-- 
predicting at points that are not data points
--
```{r}
myFit2<- function( x,y, xnew){
 degree<- 2
 fit<- lm(y~ x + I(x^2) )
 A<- fit$coefficients
 predValues<- A[1] + A[2]*xnew + A[3]*xnew^2
return( cbind( xnew, predValues) )
}
```
add a line at equally spaced sequence 
```{r}
xnew<- seq( 0,2e5,,150)
plot( AudiA4$mileage, AudiA4$price)
hold2<- myFit2(AudiA4$mileage, AudiA4$price, xnew)
lines( hold2, col="orange2", lwd=3)
```

-- 
Adding a  default choice to the xnew option. If ```xnew``` is not 
specified it will be the x data values but in sorted order.
--
```{r}
myFit3<- function( x,y, xnew= sort(x)){
 degree<- 2
 fit<- lm(y~ x + I(x^2) )
 A<- fit$coefficients
 predValues<- A[1] + A[2]*xnew + A[3]*xnew^2
return( cbind( xnew, predValues) )
}
```

--
Finally adding some extra information and returning a list with other information.
-- 
```{r}
myFit4<- function( x,y, xnew= sort(x), degree = 2 ){
# place holder in case this function is extended to other polynomials
 if( degree != 2){
 stop("Function does not handle this polynomial degree")
 }
 if( any(is.na(y))){
 warning("y has missing values!")
 }
 fit<- lm(y~ x + I(x^2) )
 A<- fit$coefficients
 predValues<- A[1] + A[2]*xnew + A[3]*xnew^2
  return( 
       list( fit = cbind( xnew, predValues),
          degree = degree,
             call= match.call()
        )
)
}
```
```{r}
mileage<-AudiA4$mileage
yTest<- AudiA4$price
yTest[1]<- NA
hold<- myFit4( mileage, yTest)
print(hold$call)
print(hold$degree)
plot( AudiA4$mileage, yTest)
lines( hold$fit, col="orange3", lwd=3)
title( hold$call)
```


